Fourmile Calibrations and Output

### 2025-04-09 params.Rmd
added model performance to parameters. need to finish formatting

```{r}
# params area weighted (only for lagSurfaceRunoff, a_rain, a_snow)
params <- read_csv('./38HRU_dec2024_updates/params_weighted_dec2024.csv') %>%
  filter(param %in% c('lagSurfaceRunoff',
                      'a_rain',
                      'a_snow')) %>% # select params in the top 8 sens that are area weighted
  rename(value = area_weighted_average,
         input_data = source) %>%
  dplyr::select(param, input_data, value)

# params unburned (above is params that needed to be area weighted bc burn/unburn)
params_unburn <- read_csv('./38HRU_dec2024_updates/params_unburn.csv') %>%
  filter(param %in% c('soilOutLPS',
                      'snow_trs',
                      'lagInterflow',
                      'baseTemp',
                      't_factor',
                      'soilLatVertLPS')) %>%
  dplyr::select(param, input_data, value)

@@

# boxplots of top 8
ggplot(param_sens, aes(x=input_data, y=value, fill=input_data)) +
  geom_boxplot(alpha=0.8,
                outlier.colour = "black") +
  facet_wrap(~param, scales = 'free_y') +
  labs(y=element_blank(),
       x=element_blank()) +
  theme_bw(base_size=20) +
  stat_compare_means(aes(group = input_data),
                     method = "wilcox.test", 
                     label = "p.signif",
                     hide.ns = T,
                     paired = F,
                     #label.y = 0,
                     #label.x = 1.4,
                     size = 7) +
  scale_fill_manual(values = c("#0047AB", "orange")) +
    theme(
    legend.position = 'none',
    #legend.justification = c(0, 1.1), 
    legend.text = element_text(size = 16),
    legend.title = element_blank(), 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    #strip.text = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))

ggsave('./figures/param_sensit_boxplots.png',
       dpi=600)

>>>>

# params area weighted (only for lagSurfaceRunoff, a_rain, a_snow)
params <- read_csv('./38HRU_dec2024_updates/params_weighted_dec2024.csv') %>%
  filter(param %in% c('lagSurfaceRunoff',
                      'a_rain',
                      'a_snow')) %>% # select params in the top 8 sens that are area weighted
  rename(value = area_weighted_average,
         input_data = source) %>%
  dplyr::select(param, input_data, year_missing, value)

# params unburned (above is params that needed to be area weighted bc burn/unburn)
params_unburn <- read_csv('./38HRU_dec2024_updates/params_unburn.csv') %>%
  filter(param %in% c('soilOutLPS',
                      'snow_trs',
                      'lagInterflow',
                      'baseTemp',
                      't_factor',
                      'soilLatVertLPS')) %>%
  dplyr::select(param, input_data, year_missing, value) %>%
  mutate(year_missing = if_else(year_missing == 'none',
                                'all',
                                year_missing))

@@

# read in pbias categories
pbias <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/pbiasCats.csv') %>%
  filter(year == c('All')) %>%
  dplyr::select(input_data, year_missing, pbias)

nse <- read_csv('/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/nseCats.csv') %>%
  filter(year == c('All')) %>%
  dplyr::select(input_data, year_missing, nse)

perf <- left_join(nse, pbias, by = c('input_data',
                                     'year_missing'))

perf <- perf %>%
  mutate(performance = case_when(
                                nse < 0.5 & abs(pbias) >= 25 ~ 'Neither',
                                nse > 0.5 & abs(pbias) >= 25 ~ 'NSE',
                                nse < 0.5 & abs(pbias) < 25 ~ '% bias',
                                nse > 0.5 & abs(pbias) < 25 ~ 'NSE + % bias')) %>%
  mutate(input_data = if_else(input_data == 'grid',
                              'Gridded', 'Station'),
         year_missing = if_else(year_missing == 'None',
                                'all', year_missing))

param_sens <- left_join(param_sens,
                        perf,
                        by = c('input_data',
                               'year_missing'))

# boxplots of top 9
ggplot(param_sens, aes(x=input_data, y=value, fill=input_data)) +
  geom_boxplot(alpha=0.8,
                outlier.shape=NA) +
  facet_wrap(~param, scales = 'free_y') +
  labs(y=element_blank(),
       x=element_blank()) +
  theme_bw(base_size=20) +
  stat_compare_means(aes(group = input_data),
                     method = "wilcox.test", 
                     label = "p.signif",
                     hide.ns = T,
                     paired = F,
                     #label.y = 0,
                     #label.x = 1.4,
                     size = 7) +
  geom_jitter(aes(shape = performance),
            position = position_jitter(width = 0.15), 
            size = 2.5, 
            alpha = 0.6,
            stroke = 1,
            color = "black") +
  scale_shape_manual(values = c(
  "% bias" = 1,
  "NSE + % bias" = 16
)) +
  scale_fill_manual(values = c("#0047AB", "orange")) +
    theme(
    legend.position = 'none',
    #legend.justification = c(0, 1.1), 
    legend.text = element_text(size = 16),
    legend.title = element_blank(), 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    #strip.text = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))

ggsave('./figures/param_sensit_boxplots_pts.png',
       dpi=600)
